Appendix

### 2022-02-11 manuscript/appendix.Rmd

Because the methods applied here: Stochastic Dynamic Programming, Iterative Forecasts assessed with Proper Scoring Rules, or Adaptive Management, are not themselves novel, the reader is referred to previous reviews and text books which can provide a thorough introduction to methods, in particular, @Marescot2013 and @Smith1981.  However, implementation of these approaches generally depends on numerical methods, for which even precise mathematical formulas or pseudo-code descriptions can be inadequate to specify unambiguously or to properly facilitate replication [@Barnes2010].  As such, this appendix includes computer code in the R language, which can also be found in the corresponding R-Markdown document in the paper's GitHub repository, <https://github.com/cboettig/bad-forecast-good-decision>.  While computer code can be more verbose and difficult to read than mathematical formulas presented here, it is also less ambiguous. Though R is already widely used among ecologists, the implementations concern mostly matrix algebra that can be translated to other languages, and the `MDPtoolbox` package used here is also implemented in MatLab [@Marescot2013; @MDPtoolbox].

@@

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ]) 
```

@@

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state]) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

@@

```{r proper_scores, message = FALSE}
# Gneiting & Raferty (2007), eq27
scoring_fn <- function(x, mu, sigma){ -(mu - x )^2 / sigma^2  - log(sigma)}

>>>>

The methods applied here -- stochastic dynamic programming, iterative forecasts assessed with proper scoring rules, and adaptive management -- are not themselves novel, thus the reader is referred to previous reviews and text books which can provide a thorough introduction to methods, in particular, @Marescot2013 and @Smith1981.  However, implementation of these approaches generally depends on numerical methods, for which even precise mathematical formulas or pseudo-code descriptions can be inadequate to specify unambiguously or to properly facilitate replication [@Barnes2010].  As such, this appendix includes computer code in the R language, which can also be found in the corresponding R-Markdown document in the paper's GitHub repository, <https://github.com/cboettig/bad-forecast-good-decision>.  While computer code can be more verbose and difficult to read than mathematical formulas presented here, it is also less ambiguous. Though R is already widely used among ecologists, the implementations concern mostly matrix algebra that can be translated to other languages, and the `MDPtoolbox` package used here is also implemented in MatLab [@Marescot2013; @MDPtoolbox].

@@

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = as.numeric( transitions[[model]][state_index, , 1]  %*% states),
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ]) 
```

@@

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = as.numeric( transitions[[model]][state_index, , action_index]  %*% states ),
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state]) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

@@

```{r proper_scores, message = FALSE}
# Gneiting & Raferty (2007), eq27
scoring_fn <- function(x, mu, sigma){ as.numeric( -(mu - x )^2 / sigma^2  - log(sigma) )}
